Extended Abstract

### 2021-05-11 paper/paper.tex
Rewrite the intro

Consider a linear \picalc{} process with a free channel variable $x$ and wherein we create a new channel $y$, use $y$ to transmit, then send $y$ away over $x$ and terminate.
Unless the payload type of $x$ is determined elsewhere, in this process the usage annotations for the channel $y$ are polymorphic, yet the output capability on $y$ requires a lower bound of multiplicity $\tone$ (since the process uses it to transmit).
While some of the constraints on $x$ and $y$ can be eagerly solved, others are non-local and need to be deferred.

We approach this problem by breaking down typing inference into the collection of constraints \autoref{inference} and their (partial) resolution \autoref{constraint-resolution}.
We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices \autoref{syntax}.
The syntax for processes contains a small expression syntax that handles composite sum and product types.
On top of the well-scoped syntax we define a standard linear type system using context-splits \autoref{type-system}.

Constraints are defined on types and usage annotations alike \autoref{constraints}.
We use a common metavariable context for both types and usage annotations.
The metavariable context is kinded, ans substitution is defined so that it respects kindedness.
In \autoref{inference} constraints are collected bottom-up, making type-inference easily parallelisable.
Context splits are solved by creating a fresh extra context and constraints that \emph{merge} the three.
We prove a soundness theorem and postulate a completeness theorem, both allowing for constraint solving to be deferred \autoref{inference-theorems}.

Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of \cite{McBride03}.
Constraints of the form $\eqconstr{x}{y}{z}$ are solved recursively until a base case is reached, when they are either admitted as valid or deferred.
We postulate soundness and completeness theorems and show that only those constraints without a unique solution are deferred.

@@

Nothing prevents us from returning the original constraints as output.
We therefore promise that every constraint we return is either currently
unsolvable because we have not enough information, or is unsatisfiable
altogether.
\begin{flalign*}
&\func{solve}~cs_1 \equiv (\sigma, cs_2) \to \forall c \in cs_2. ~ \type{IsDeferred}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
\end{flalign*}
where
\begin{flalign*}
& \type{IsUnsat}~c \triangleq \forall \sigma. ~ \neg \interpr{\subst{\sigma}{c}} &&
\end{flalign*}

\paragraph{Instantiation.}

\begin{flalign*}
& \forall c. ~ \type{IsDeferred}~c \to \exists \sigma. \interpr{\subst{\sigma}{c}} &&
\end{flalign*}

>>>>

Consider a linear \picalc{} process with a free channel variable $x$ and wherein we create a new channel $y$, use $y$ to transmit, then send $y$ away over $x$ and terminate.
Unless the payload type of $x$ is determined elsewhere, the usage annotations for the channel $y$ are polymorphic, and the output capability on $y$ requires a lower bound of multiplicity $\tone$ (since the process uses it to transmit).
This lower bound is a constraint that cannot be solved locally: it needs to be deferred until it is further constrained.

We break down typing inference into the collection of constraints (\autoref{inference}) and their (partial) resolution (\autoref{constraint-resolution}).
We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices (\autoref{syntax}).
The syntax for processes contains a small expression language that handles composite sum and product types.
On top of the well-scoped syntax we define a standard linear type system that uses context-splits (\autoref{type-system}).

Types and usage annotations contain kinded metavariables (holes) part of a common metavariable context.
Substitution respects kindedness, and constraints are uniformly defined over both types and usage annotations.
Constraints are collected bottom-up (\autoref{inference}), which makes type-inference easily parallelisable.
Context splits are solved by creating a fresh extra context and constraints that \emph{merge} the three.
We prove a soundness theorem and postulate a completeness theorem, both use deferred constraints (\autoref{inference-theorems}).

Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of \cite{McBride03}.
Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, when they are either admitted as valid or deferred.
We postulate soundness and completeness theorems and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution-theorems}).

@@

To keep us from returning the original constraints as output we show that every constraint we return is either unsatisfiable or not uniquely constrained and has to be deferred.
altogether.
\begin{flalign*}
&\func{solve}~cs_1 \equiv (\sigma, cs_2) \to \forall c \in cs_2. ~ \type{IsDeferred}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
\end{flalign*}
where
\begin{flalign*}
& \type{IsUnsat}~c \triangleq \forall \sigma. ~ \neg \interpr{\subst{\sigma}{c}} && \\
& \type{IsNotUnique}~c \triangleq \exists \sigma_1. \interpr{\subst{\sigma_1}{c}} \times \exists \sigma_2. ~ \interpr{\subst{\sigma_2}{c}} \times \sigma_1 \not\doteq \sigma_2 &&
\end{flalign*}
